---
title: "Exploring Income Inequality Dynamics in Canada: A Geospatial and Temporal Analysis"
subtitle: "Insights from 1976 to 2021"
author: 
  - Sirui Tan
thanks: "Code and data are available at: https://github.com/siru1366/income.git."
date: today
date-format: long
abstract: "This paper investigates income inequality in Canada from 1976 to 2021 using data from various surveys. Through exploratory data analysis and regression modeling, we uncover increasing income inequality trends over time and variations across different provinces. Our findings reveal nuanced relationships between temporal trends and regional disparities, and income distribution dynamics. Through rigorous analysis, we uncover two key findings: a consistent rise in income inequality over time, significant regional disparities across provinces. Understanding these dynamics is crucial for policymakers to develop targeted strategies aimed at mitigating income inequality and promoting social and economic equity in Canada."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
library(broom.mixed)
library(beepr)
library(broom)
library(knitr)
library(purrr)
library(rstanarm)
library(tidyverse)
library(bayesplot)
library(coda)
library(kableExtra)
library(dplyr)
library(tidyr)
library(arrow)
library(modelsummary)


```

# Introduction
Income inequality is a pressing socioeconomic issue that has garnered significant attention worldwide, with Canada being no exception. Over the past few decades, Canada has witnessed a widening gap in per capita after-tax income, raising concerns about economic disparities and social justice. Understanding the drivers behind this trend is crucial for policymakers and researchers alike, as it provides insights into the factors influencing income distribution and informs strategies to mitigate inequality.

This paper aims to delve into the intricate dynamics of income inequality within Canada, focusing on the period from 1976 to 2021. Leveraging a comprehensive dataset spanning multiple surveys, including the Survey of Consumer Finances (SCF), the Survey of Labour and Income Dynamics (SLID), and the Canadian Income Survey (CIS), we embark on a rigorous exploration of income trends across different provinces and over time. By examining key variables such as year, geographical location, income decile, and income range, we seek to uncover patterns and nuances in income distribution dynamics.

Despite numerous studies on income inequality in Canada, there remains a clear gap in understanding the specific drivers and mechanisms contributing to the widening income gap observed over the past four decades. To address this gap, we undertake a multifaceted analysis, encompassing exploratory data analysis and multiple linear regression modeling. Through this approach, we aim to elucidate the role of various factors, including temporal trends, regional disparities, in shaping income distribution patterns.

Temporal trends reveal a consistent increase in income inequality from 1976 to 2021. Geographical disparities highlight varying levels of income inequality across provinces, with Atlantic provinces and Quebec showing higher inequality compared to others.Overall, these findings emphasize the multifaceted nature of income inequality dynamics in Canada, urging policymakers to consider both temporal trends and regional disparities when addressing this pressing socioeconomic issue.

The remainder of this paper is structured as follows. @sec-data introduces the data used for analysis and findings, including visualizations of the variables of interest, @sec-model
proposes a straightforward linear regression model to examine and forecasts the effects of time (year) and geographical location (province) on income inequality in Canada.In @sec-results, we display the interpretations of the models alongside other findings from analyzing the data. @sec-discussion provides a discussion on the implications of the findings as well as the weaknesses of this paper and its next steps for further study on this subject.

## Estimand
The estimand of this study is to quantify the relationship between time (year) and income inequality in Canada. Specifically, we aim to estimate the effect of each unit change in the year on income inequality, holding other variables constant. This involves analyzing how income inequality trends have evolved over the period from 1976 to 2021, capturing any systematic changes over time.

Additionally, we seek to understand the impact of geographical location (province) on income inequality and how this interacts with temporal trends. By examining the coefficients associated with both year and province variables in our regression models, we aim to discern the average differences in income inequality between different provinces, while also exploring how these differences may have changed over time.

Overall, our estimand focuses on uncovering the nuanced relationship between time, geographical location, and income inequality in Canada, providing insights into the factors influencing income distribution dynamics over the study period.

# Data {#sec-data}

## Data Source

Upper income limit, income share and average income by economic family type and income decile

## Data Measurment

The estimates for Upper Income Limit, After-Tax Income by Economic and Income Decile in Canada are based on data from various surveys spanning different periods. From 1976 to 1992, data was collected from the Survey of Consumer Finances (SCF). For the period from 1993 to 1997, a combination of SCF and the Survey of Labour and Income Dynamics (SLID) was used. Subsequently, from 1998 to 2011, data solely from SLID was utilized. Starting from 2012, the Canadian Income Survey (CIS) became the primary data source. More details on these surveys and their revisions can be found in publications by Statistics Canada, including "Revisions to 2006 to 2011 Income Data" (2015) by Statistics Canada and other related papers by Cotton (2000) and Lathe (2005).

The Canadian Income Survey (CIS) aims to provide insights into the income and its sources of Canadians, alongside their individual and household characteristics. It combines data from the Labour Force Survey (LFS) and tax records.

Estimates from the Survey of Consumer Finances cover individuals aged 15 years and over, whereas estimates from SLID and CIS include individuals aged 16 years and over.

The CIS introduced improvements in methodology and data processing, notably from the 2021 reference year onwards. It transitioned to using the Administrative Personal Income Masterfile, incorporating data from both T1 tax returns and associated tax slips. Prior to this, only T1 tax returns were used. These enhancements, including updates to weighting methodology, aim to enhance data quality while minimally impacting key estimates and trends.

The CIS is a sample survey with a cross-sectional design, administered to a sub-sample of LFS respondents. The LFS employs a rotating panel sample design, with selected dwellings remaining in the sample for six consecutive months in the provinces and for two years in the territories. Rotation groups from the LFS are utilized for the CIS sample, with approximately 55,000 households included in the CIS sample for 2021.

Data cleaning and analysis were conducted using the open-source statistical programming language `R` [@citeR], leveraging functionalities from the `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tibble` [@tibble],`stringr`  [@stringr], `haven` [@haven], `janitor` [@janitor], `knitr` [@knitr].

## Variables of Interest

```{r dataextract, echo=FALSE, message=FALSE}
# create a table showing the first 20 rows of the cleaned data
income <<- read_parquet(file = here::here("data/analysis_data/cleaned_income_data.parquet"))
income %>%
  slice(1:20) %>%
  kable(
    caption = "Extracting the first ten rows from the Income data",
col.names=c("Year", "Geographical location", "Income decile", "Income", "Highest-to-Lowest Average Income Ratio", "income range"),
linesep = "", digits = 1, booktabs=TRUE) %>%
  kable_styling(latex_options = "HOLD_position", font_size = 10)
```


1.  **Year**: The dataset spans from 1976 to 2021, covering a wide temporal range that allows for the exploration of long-term income trends and changes over time. This extended period facilitates the analysis of income dynamics across different decades, offering insights into patterns and shifts in income distribution.

2.  **Geographical location**: The dataset includes data from all provinces in Canada, providing a comprehensive view of income distribution dynamics across the entire country. This geographic diversity enables the examination of regional variations and disparities in income levels and inequality measures.

3.  **Income decile**: All the units of the population, whether economic families or persons not in an economic family, are ranked from lowest to highest by the value of their income of a specified income concept. Then, the ranked population is divided into ten groups of equal numbers of units, called deciles. Individuals are grouped into ten equal categories based on their income levels within each province and year. These income deciles offer a systematic framework for analyzing income distribution patterns, allowing for comparisons across different segments of the population and over time.

4.  **Income**: The dataset contains information on average income levels for each income decile within every province and year. These income values offer insights into the economic well-being of individuals and households across various regions of Canada, enabling the assessment of income disparities and socioeconomic conditions over the years.

5.  **Highest-to-Lowest Average Income Ratio**: Calculated for each province and year, this ratio measures the degree of income inequality within the population. It represents the ratio of the average income of the highest income decile to that of the lowest income decile within a specific province and year, providing a standardized measure of income concentration or inequality across different regions and time periods in Canada.

6.  **Income Range**: The income range data provides detailed information about income distribution across different years, geographical locations, and income deciles in Canada. For instance, in 1976, the total income for Canada was 63,300, with the highest-to-lowest average income ratio being 16.8, indicating a significant income gap between the highest and lowest deciles. The income range for the lowest decile in Canada was 9,400, while the income for the highest decile was $157,500.

We focus on both the Highest-to-Lowest Average Income Ratio and income range because they provide complementary insights into income inequality and distribution within a population. The Highest-to-Lowest Average Income Ratio offers a standardized measure of income concentration or inequality by comparing the average income of the highest income decile to that of the lowest income decile. This ratio allows us to understand the extent of income disparity within a specific group or region over time.

On the other hand, the income range provides information about the actual income levels across different deciles within the population. By examining the range of incomes from the lowest to the highest decile, we can identify the magnitude of income discrepancies and the economic well-being of individuals or households at various income levels. This data helps in understanding the distribution of wealth and assessing the socioeconomic conditions within a society.

## Data Visualization


```{r}
#| label: fig-bills
#| fig-cap: Trend of Average After-Tax Income Across Income Deciles in Canada in 1976 to 2021
#| echo: false
#| warning: false
income3 <- read_parquet(file = here::here("data/analysis_data/cleaned_income_data.parquet"),show_col_types = FALSE)

income3 %>%
  filter(
         `Income decile` == "Lowest decile" | `Income decile` == "Total deciles"| `Income decile` == "Highest decile",
         `Geographical_Location` == "Canada") %>%
  ggplot(aes(x=Year, y=Income, colour = `Income decile`))+geom_line()+theme_minimal()+ylab("Average after-tax income")
  

```

As depicted in the figure by @fig-bills, the average after-tax income for economic families and unattached individuals in Canada has demonstrated a consistent upward trajectory from 1976 to 2021. During this period, the average after-tax income in Canada increased from approximately \$63,300 in 1976 to \$83,100 in 2021. Notably, after 1997, the highest decile of average after-tax income for economic families and unattached individuals experienced a notably accelerated growth rate compared to the overall average. Conversely, during the same period, the lowest decile of average after-tax income for economic families and unattached individuals exhibited sluggish growth.


```{r fig.width=6, fig.height=10}
#| label: fig-planes
#| fig-cap: Analyzing 1976-2021 Income Disparity Trends in Canada by Highest-to-Lowest Average Income Ratio
#| echo: false
#| warning: false
#| message: false
income3 <- read_parquet(file = here::here("data/analysis_data/cleaned_income_data.parquet"),show_col_types = FALSE)
income3 %>%
  filter(
         `Income decile` == "Total deciles",
         `Geographical_Location` != "Canada") %>%
  ggplot(aes(x=Year, y=`Highest-to-Lowest_Average_Income_Ratio`, colour = `Geographical_Location`))+geom_point()+theme_minimal()+ylab("Average after-tax income range")
```

@fig-planes, the Highest-to-Lowest Average Income Ratio across Canadian provinces has shown a general upward trend, following a decline from 1980 to 1990. However, each province exhibits distinct performance. Notably, British Columbia's ratio surpasses that of other provinces. Given that this ratio reflects relative disparities and is heavily influenced by the lowest decile of average after-tax income for economic families and unattached individuals, deviations from the norm are evident in certain data points.
```{r}
#| label: fig-planes2
#| fig-cap: Analyzing 1976-2021 Income Disparity Trends in Canada by Average after-tax Income Range
#| echo: false
#| warning: false
#| message: false
income4 <- read_parquet(file = here::here("data/analysis_data/cleaned_income_data.parquet"),show_col_types = FALSE)
income4 %>%
  filter(
         `Income decile` == "Total deciles",
         `Geographical_Location` != "Canada") %>%
  ggplot(aes(x=Year, y=`Income_Range`, colour = `Geographical_Location`))+geom_point()+theme_minimal()+ylab("Average after-tax income range")
```


The @fig-planes2 illustrates another metric of income inequality: the variance between the lowest and highest deciles of average after-tax income for economic families and unattached individuals. This disparity is more pronounced compared to the Average after-tax Income Range. Each province displays distinct performance in this regard. Notably, Prince Edward Island's value remains lower than that of other provinces, suggesting relatively greater income equality within this region.

# Model {#sec-model}

Through our exploratory data analysis, we uncovered compelling evidence suggesting a correlation between both the year and province variables and the level of income inequality across Canada. Notably, our observations indicate a linear relationship, with figures consistently showing upward trends over time. Given these findings, we are prompted to delve deeper into the analysis and forecast future trends in income inequality. To achieve this, we will develop a multiple linear regression model.

We aim to investigate the association between the highest-to-lowest average income ratio of average after-tax income to the lowest decile of average after-tax income and the after-tax income gap, considering the factors of year, province. To accomplish this, we developed two distinct multiple linear regression models.


## Model set-up

\begin{align}
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 \times \mbox{Year}_i + \beta_2 \times \mbox{Geographical Location}_i\\
\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\label{eq-model1}
\end{align}

In Model (\ref{eq-model1})is specified as follows:

-   $Y_{i}$ represents the highest-to-Lowest average income ratio of average after-tax income to the lowest decile of average after-tax income of Canada in $i^{th}$ year and province $i$.

-   $\beta_{0}$ is the coefficient for intercept.

-   $\beta_{1}$ represents the coefficient for the variable $Year_{i}$, capturing the effect of time (year) on income inequality. A positive coefficient suggests that income inequality tends to increase over time, while a negative coefficient suggests a decrease.

-   $\beta_{2}$ represents the coefficient for the variable $Province_{j}$, indicating the effect of province on income inequality. It accounts for regional differences in income distribution within Canada.

\begin{align}
y_i|\mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \\
\mu_i &= \beta_0 + \beta_1 \times \text{Year}_i + \beta_2 \times \text{Geographical Location}_i\\
\beta_0 &\sim \text{Normal}(0, 2.5) \\
\beta_1 &\sim \text{Normal}(0, 2.5) \\
\beta_2 &\sim \text{Normal}(0, 2.5) \\
\sigma &\sim \text{Exponential}(1)
\label{eq-model2}
\end{align}

In Model (\ref{eq-model2}):


-   $Y_{i}$ represents the income range for a particular combination of year ($Year_{i}$) and geographical location ($Province_{j}$).

-  $\beta_{0}$ is the intercept term, indicating the expected value of the income range when all other predictors are zero.

- $\beta_{1}$ is the coefficient associated with the "Year" variable ($Year_{i}$), representing the effect of each unit change in the year on the income range, holding other variables constant.

- $\beta_{2}$ is the coefficient associated with the "Geographical Location" variable (\(Province_{j}\)), representing the average difference in income range between different provinces, holding the year constant.


We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

## Model justification

The first model (Equation 1) aims to capture the relationship between the response variable, $y_i$, representing the highest-to-lowest average income ratio of average after-tax income to the lowest decile of average after-tax income of Canada in the $i^{th}$ year and province $j$, and the predictor variables, $\text{Year}_i$ and $\text{Geographical Location}_i$.

The mean, $\mu_i$, is modeled as a linear combination of the intercept, $\beta_0$, and the coefficients for the predictor variables, $\beta_1$ for $\text{Year}_i$ and $\beta_2$ for $\text{Geographical Location}_i$. This allows us to assess the impact of time (year) and province on income inequality.

The choice of priors for the coefficients, $\beta_0$, $\beta_1$, and $\beta_2$, as Normal distributions with mean 0 and a standard deviation of 2.5 reflects a weak prior assumption, indicating that we have no strong prior beliefs about the magnitude or direction of their effects.

The standard deviation parameter, $\sigma$, is modeled using an Exponential distribution with a rate parameter of 1. This distribution allows for uncertainty in the variability of the response variable around the mean.

Overall, this model provides a flexible framework for analyzing the relationship between income inequality and time, as well as geographical location, while incorporating uncertainty through the specification of appropriate prior distributions.


Model (\ref{eq-model2}) is formulated to explore the relationship between the income range ($Y_{ij}$) and the predictor variables, Year ($Year_{i}$), and Geographical Location ($Province_{j}$).

$Y_{ij}$ represents the income range for a specific combination of year ($Year_{i}$) and geographical location ($Province_{i}$). This formulation allows us to examine how income ranges vary across different years and provinces.

$\beta_{0}$ serves as the intercept term in the model, representing the expected value of the income range when all other predictors are zero. It provides a baseline reference point for comparison.

$\beta_{1}$ is the coefficient associated with the "Year" variable ($Year_{i}$). This coefficient quantifies the effect of each unit change in the year on the income range, while holding all other variables constant. A positive coefficient suggests an increase in the income range over time, while a negative coefficient indicates a decrease.

$\beta_{2}$ is the coefficient associated with the "Geographical Location" variable ($Province_{i}$). This coefficient captures the average difference in income range between different provinces, assuming the year remains constant. It allows us to assess the impact of geographical location on income disparities.

Model 2 provides a structured framework for analyzing the factors influencing income range variations across different years and geographical locations. By examining the effects of both temporal and spatial factors, the model helps to elucidate patterns of income distribution and disparities within the studied context.



# Results {#sec-results}



```{r figure5, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Difference between lowest and highest average after-tax income in the 10 provinces of Canada between 1976 and 2021", fig.width=10, fig.height=3}
# plot the graph with the regression lines
income <- read_csv(here::here("data/analysis_data/cleaned_income_data.csv"))
income <- income %>%
  filter(
    `Income decile` == "Total deciles",
    `Geographical Location` != "Canada")
income %>%
  ggplot(aes(x=Year, y=`Income Range`, colour = `Geographical Location`))+geom_point()+theme_minimal()+ylab("Highest-to-Lowest Average Income Ratio ")+geom_smooth(method='lm')
```





```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
library(rstanarm)

first_model <- 
  readRDS(file = here::here("models/model2.rds"))


second_model<- 
  readRDS(file = here::here("models/model.rds"))

modelsummary(
  list(
    "Model 1" = first_model,
    "Model 2" = second_model
  ), statistic = "mad",
  fmt = 2
)

```




```{r}
#| label: tbl-model1
#| tbl-cap: "Summary results for two models"
#| echo: false
#| eval: true
#| warning: false
#| message: false
first_model <- 
  readRDS(file = here::here("models/model2.rds"))


second_model<- 
  readRDS(file = here::here("models/model.rds"))

modelsummary::modelsummary(
  list(
    "Model 1" = first_model,
    "Model 2" = second_model
  ), statistic = "mad",
  fmt = 2
)


```

@tbl-model1

The regression results are as follows:

The results of Model 1 reveal several important findings regarding the relationship between the predictor variables and the highest-to-lowest average income ratio.

Firstly, the intercept term is calculated to be -98.40, indicating the estimated average ratio when all other predictor variables are zero. This intercept provides a baseline for comparison against the effects of the other predictors.

The coefficient for the variable "Year" is estimated to be 0.06. This positive coefficient suggests a slight increase in the highest-to-lowest average income ratio over time, although the effect appears to be relatively small.

Furthermore, the coefficients for various provinces show differing effects on the income ratio compared to the reference category. For instance, provinces such as Alberta and British Columbia exhibit positive coefficients of 3.91 and 8.28, respectively, indicating higher average income ratios compared to the reference category. Conversely, provinces like Newfoundland and Labrador, Prince Edward Island, and New Brunswick demonstrate negative coefficients, suggesting lower average income ratios in these regions.

Overall, the model's R-squared value of 0.149 indicates that approximately 14.9% of the variability in the highest-to-lowest average income ratio is explained by the predictor variables included in the model. Additionally, the adjusted R-squared value of 0.117 suggests that the model may slightly overestimate the true explanatory power due to the inclusion of additional predictors.

The log likelihood of -1934.210 indicates the goodness-of-fit of the model, with lower values indicating a better fit. Furthermore, the root mean squared error (RMSE) of 8.25 provides a measure of the model's predictive accuracy, with smaller values indicating better predictive performance.

In summary, Model 1 provides valuable insights into the factors influencing the highest-to-lowest average income ratio in Canada, highlighting the effects of time (Year) and geographical location (provinces) on income inequality. These findings contribute to a better understanding of regional disparities in income distribution and can inform policy decisions aimed at reducing inequality across different provinces.



The results of the model2 reveal several significant findings regarding the relationship between the response variable and the predictor variables. Firstly, the intercept term is estimated to be -3309430.66, indicating the expected value of the response variable when all other predictors are zero. This intercept provides a reference point for comparison and interpretation of the effects of the other predictor variables.

The coefficient for the variable "Year" is calculated to be 1730.57. This positive coefficient suggests that there is an average increase of 1730.57 units in the response variable for each unit increase in the year, holding all other variables constant. This implies that over time, there is a general upward trend in the response variable.

Furthermore, the coefficients associated with the different geographical locations (provinces) indicate the average difference in the response variable compared to the reference category. For instance, provinces like Ontario, Prairie Provinces, Alberta, and British Columbia exhibit positive coefficients, suggesting higher values of the response variable compared to the reference category. Conversely, provinces like Atlantic Provinces, Newfoundland and Labrador, Prince Edward Island, Nova Scotia, New Brunswick, Quebec, and Manitoba demonstrate negative coefficients, indicating lower values of the response variable in these regions.

Overall, the model demonstrates a good fit to the data, as indicated by the high values of R-squared (0.794) and adjusted R-squared (0.787). These metrics suggest that approximately 79.4% of the variability in the response variable is explained by the predictor variables included in the model. Additionally, the log likelihood value of -6054.760 indicates a good fit of the model to the data.

In summary, the results provide valuable insights into the factors influencing the response variable, as well as the overall fit and predictive performance of the model. These findings contribute to a better understanding of the relationships between the predictor variables and the response variable, which can inform decision-making and policy development in relevant contexts.




# Discussion {#sec-discussion}

## Findings {#sec-first-point}

This study examined income inequality in Canada from 1976 to 2021 using various survey data. Through data analysis and regression modeling, we found that income inequality has consistently increased over time and varies among provinces. Our key findings include a continual rise in income inequality, significant regional differences, and complex interactions between temporal trends and income distribution.

Firstly, our analysis revealed a clear upward trend in income inequality over the study period, as indicated by the positive coefficient for the "Year" variable in our regression models. This suggests that income inequality has steadily worsened in Canada over the years. Secondly, we found notable differences in income inequality levels across provinces, with some regions experiencing higher disparities than others. 

In summary, our study provides valuable insights into income inequality dynamics in Canada, underscoring the need for policymakers to consider both temporal trends and regional differences when devising strategies to mitigate income inequality and promote economic equity.



## Analyzing Temporal Trends in Income Inequality in Canada

The widening income gap in Canada, particularly between the top and bottom 10%, has persisted since 1976 due to various socioeconomic factors.

Firstly, changes in economic structure and globalization have favored certain industries and skill sets, leading to increased demand and higher wages for individuals with specialized education or in high-demand sectors. This has disproportionately benefited the top earners, contributing to their rising after-tax income @Autor2014.

Additionally, government policies and tax reforms over the years have also played a role. Tax cuts and preferential treatment for high-income individuals and corporations may have further widened the income gap by allowing the wealthy to accumulate more wealth while providing limited benefits to low-income earners @Piketty2014.

Furthermore, technological advancements and automation have transformed the labor market, leading to job polarization. While high-skilled jobs requiring advanced education or technical expertise have seen wage growth, low-skilled jobs have faced stagnation or even decline in wages. This has perpetuated income inequality as those at the top benefit from technological advancements, while those at the bottom struggle to keep up @Acemoglu2019.

Moreover, systemic issues such as disparities in access to education, healthcare, and opportunities for upward mobility have also contributed to the widening income gap. Structural inequalities embedded within society perpetuate cycles of poverty, making it difficult for individuals from lower-income backgrounds to break free from economic hardship @Chetty2020.

Overall, the combination of economic, policy, technological, and social factors has led to the widening income gap in Canada, highlighting the pressing need for comprehensive strategies to address income inequality and promote economic inclusion and equity for all segments of society.

## Exploring Factors Contributing to Varied Income Inequality Across Canadian Provinces
The disparity in income inequality between provinces in Canada can be attributed to a multitude of factors, including economic structure, industry composition, demographics, and government policies @smith2019. While some provinces like British Columbia exhibit higher ratios of the highest income to the lowest income and wider income gaps, others like Prince Edward Island (PEI) demonstrate comparatively lower values @jones2020. Several key reasons can elucidate why PEI may have lower income inequality metrics:

Economic Structure: PEI's economic structure differs significantly from provinces like British Columbia. As a smaller province with a predominantly rural economy, PEI relies heavily on industries such as agriculture, fisheries, and tourism @canadaEconomicReport. These sectors often have more equitable income distributions compared to industries like finance or technology, which contribute to higher income inequality in provinces with more diversified economies.

Population Size and Density: PEI's smaller population size and lower population density relative to provinces like British Columbia can influence income inequality. With a smaller labor force and fewer high-income earners, PEI may experience less income disparity @canadaCensus. Additionally, the close-knit communities and social cohesion in smaller provinces like PEI can contribute to a more equitable distribution of resources and opportunities @peiCommunityReport.

Government Policies: Provincial government policies and social programs play a crucial role in mitigating income inequality. PEI's government may implement policies focused on social welfare, affordable housing, and income support programs to address economic disparities and ensure a more equitable distribution of wealth @peiGovernmentPolicy. These interventions can help reduce income inequality by providing assistance to low-income individuals and families.

Industry Composition: The types of industries dominant in a province can impact income distribution. In PEI, industries like agriculture and fisheries may offer more uniform income levels across workers, resulting in lower income inequality @peiEconomicReview. Conversely, provinces with a higher concentration of high-paying industries, such as technology or finance, may experience wider income disparities.

Cost of Living: Variations in the cost of living between provinces can influence income inequality. Provinces with lower costs of living, like PEI, may have lower income inequality as housing, healthcare, and other essential expenses consume a smaller portion of residents' incomes @canadaCostOfLiving. In contrast, provinces with higher costs of living may experience greater income disparities as a larger share of income is allocated to basic necessities.

In summary, Prince Edward Island's lower values in terms of income inequality metrics can be attributed to its unique economic structure, smaller population size, government policies, industry composition, and cost of living compared to provinces like British Columbia. These factors collectively contribute to a more equitable distribution of income within the province.


## Weaknesses and next steps

One weakness of this study is the potential for omitted variable bias. Despite our efforts to include key variables such as year and geographical location in our regression models, there may still be unobserved factors influencing income inequality that are not accounted for in our analysis. These omitted variables could lead to biased estimates of the coefficients and affect the accuracy of our findings.

Another weakness is the reliance on secondary data sources, such as surveys like the Survey of Consumer Finances (SCF), the Survey of Labour and Income Dynamics (SLID), and the Canadian Income Survey (CIS). While these datasets provide valuable information on income distribution, they may suffer from measurement error or sampling biases that could impact the reliability of our results. Additionally, the scope and coverage of these surveys may vary over time, which could introduce inconsistencies in our analysis.

Studying income inequality across occupations represents a crucial next step in understanding the broader socio-economic landscape. This research avenue involves analyzing wage differentials, earnings distributions, and the role of education and skills in driving income variation within specific job categories. Additionally, exploring intersectional dynamics, such as the impact of gender, race, and ethnicity on income inequality within occupations, is essential. Longitudinal studies tracking changes in income differentials over time can provide insights into evolving trends and challenges. By uncovering disparities and systemic biases within the labor market, this research can inform evidence-based policy interventions aimed at promoting economic equity and opportunity for all workers.

\newpage

# References

